---
title: "Data Visualization"
author: "Camila Vargas Poulsen"
date: "2023-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(leaflet)
library(DT)
```

## Read Data

```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

esc <- tryCatch(
    read.csv("data/escapement.csv"),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"))
        return(esc)
    }
)

head(esc)
glimpse(esc)
```


## Summary Data
Calculating annual escapements by species and region.

```{r}
annual_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(Species, SASAP.Region, Year) %>% 
  summarise(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))

```


## Plot Static Figures

Basic `ggplot` figure.

```{r}
ggplot(annual_esc,
       aes(x = Species,
           y = escapement))+
  geom_col()
```

Basic plot with blue bars.

```{r}
ggplot(annual_esc,
       aes(x = Species,
           y = escapement))+
  geom_col(fill = "blue")
```


Plotting escapement for each species by region

```{r}
ggplot(annual_esc,
       aes(x = Species,
           y = escapement,
           fill = SASAP.Region))+
  geom_col()
  
```

## `ggplot` and the pipe operator

Annual escapement from Kodiak by species

```{r}
annual_esc %>% 
  filter(SASAP.Region == "Kodiak") %>% 
  ggplot(aes(x = Year,
             y = escapement,
             color = Species))+
  geom_line()+
  geom_point()
  
```


## Customize our plot using theme_

```{r}
kodiak_esc <- annual_esc %>% 
  filter(SASAP.Region == "Kodiak")
```

```{r}
ggplot(kodiak_esc,
       aes(x = Year,
           y = escapement,
           color = Species))+
  geom_line()+
  geom_point()+
  ylab("Escapement")+
  ggtitle("Kodiak Salmon Escapement")+
  theme_bw()+
  theme(legend.position = "bottom",
        legend.title = element_blank())

```


## Saving my custom theme

```{r}
my_theme <- theme_bw()+
  theme(legend.position = "bottom",
        legend.title = element_blank())
```

```{r}
ggplot(kodiak_esc,
       aes(x = Year,
           y = escapement,
           color = Species))+
  geom_line()+
  geom_point()+
  scale_y_continuous(labels = comma)+
  ylab("Escapement")+
  ggtitle("Kodiak Salmon Escapement")+
  my_theme

```


## Saving my plot

```{r}
ggsave("kodiak_esc.png", width = 12, height = 3, units = "in")
```

## Crating multiple plots
Plotting escapment thoughout years by region for each species

```{r}
ggplot(annual_esc,
       aes(x = Year,
           y = escapement,
           color = Species))+
  geom_line()+
  geom_point()+
  scale_y_continuous(labels = comma)+
  facet_wrap(~SASAP.Region,
             scales = "free_y",
             ncol = 2)+
  ylab("Escapement")+
  theme_classic()
  my_theme
```

## Creating interactive tables with `DT`

Data frame with location information
```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()
```


Interactive table

```{r}
datatable(locations)
```


## Interactive maps using `leaflet`
Map with locations with OpensStreetMap

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lng = ~Longitude,
             lat = ~Latitude,
             popup = ~Location)
```

Map locations with a Web Map Service (WMS) for General Bathymetric map of the Ocean (GEBCO)

```{r}

leaflet(locations) %>% 
  addWMSTiles("https://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv?request=getmap&service=wms&BBOX=-90,-180,90,360&crs=EPSG:4326&format=image/jpeg&layers=gebco_latest&width=1200&height=600&version=1.3.0",
              layers = 'GEBCO_LATEST',
              attribution = "Imagery reproduced from the GEBCO_2022 Grid, WMS 1.3.0 GetMap, www.gebco.net") %>%
  addCircleMarkers(lng = ~Longitude,
                   lat = ~Latitude,
                   popup = ~ Location,
                   radius = 5,
                   # set fill properties
                   fillColor = "salmon",
                   fillOpacity = 1,
                   # set stroke properties
                   stroke = T,
                   weight = 0.5,
                   color = "white",
                   opacity = 1)

```


