---
title: "Practice Cleaning and Manipulating Data"
author: "Camila Vargas Poulsen"
date: "2023-03-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Inroduction
Work with Lanctot & Saalfeld data on Utqiagvik shorebird breeding ecology study to answer the following questions.

### Research Questions

1. What species of predator is the most abundant, and has this changed through time?

2. Does the number of eggs predated increase with the total number of predators for all species laying nests?

## Load Libraries

```{r}
library(rvest)
library(readr)
library(dplyr)
library(janitor)

```

## Scraping Species code from the web

```{r}
webpage <- read_html("https://www.pwrc.usgs.gov/bbl/manual/speclist.cfm")

tbls <- html_nodes(webpage, "table") %>% 
    html_table(fill = TRUE)

species <- tbls[[1]] %>% 
    clean_names() %>% 
    select(alpha_code, common_name) %>% 
    mutate(alpha_code = tolower(alpha_code))
```

## Read data

```{r}
pred <- read_csv("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A9ffec04c-7e2d-41dd-9e88-b6c2e8c4375e")


nests <- read_csv("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A982bd2fc-4edf-4da7-96ef-0d11b853102d")
```

## Defining function

```{r}
#' Function to add common name to data.frame according to the BBL list of species codes

#' @param df A data frame containing BBL species codes in column `species`
#' @param species A data frame defining BBL species codes with columns `alpha_code` and `common_name`
#' @return A data frame with original data df, plus the common name of species

assign_species_name <- function(df, species){
    
    return_df <- left_join(df, species, by = c("species" = "alpha_code"))
    
    return(return_df)
}

```

## Question 1: What species of predator is the most abundant and has this changed through time?

Calculate the total annual count for each species of predator with its common name. 

```{r}

pred_species <- assign_species_name(pred, species) %>% 
  group_by(year, common_name) %>% 
  summarise(pred_count = sum(count, na.rm = T), .groups = "drop") %>% 
  filter(!is.na(common_name)) %>%
  arrange(desc(pred_count))

```

Plotting the result to see how it changes through time.

```{r}
ggplot(pred_species, 
       aes(x = year, 
           y = pred_count, 
           color = common_name)) +
  geom_line() +
  geom_point() +
  labs(x = "Year", 
       y = "Number of Predators", 
       color = "Species") +
  theme_bw()

```

## Question 2: Does the number of eggs predated increase with the total number of predators for all species laying nests?

Calculating number of eggs per species predated by year.

```{r}
nests_species <- assign_species_name(nests, species) %>% 
  group_by(common_name, year) %>% 
  summarise(total_predated = sum(number_eggs_predated, na.rm = T),
            .groups = "drop") %>% 
  filter(!is.na(common_name))
```

Calculation the total number of predators across all species.

```{r}
pred_total <- pred_species %>% 
    group_by(year) %>% 
    summarise(pred_count = sum(pred_count, na.rm = T), 
              .groups = "drop")
```

Joing nest predation with total predation data

```{r}
nest_pred <- left_join(nests_species, pred_total, by = "year")
```

Plot total number of eggs predated over total predators and facet by species.

```{r}
ggplot(nest_pred, 
       aes(x = pred_count, y
           = total_predated)) +
    geom_point() +
    facet_wrap(~common_name, 
               scales = "free_y", 
               ncol = 2) +
    labs(x = "Number of Predators", 
         y = "Number of Eggs Predated") +
    theme_bw()
```


