---
title: "Geospatial Analysis with sf"
author: "Camila Vargas Poulsen"
date: "2023-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load libraries
```{r loadLibraries}

library(sf)
library(ggplot2)
library(ggmap)
library(scales)
library(leaflet)
library(dplyr)

```

## Read data

```{r readData}
ak_region <- read_sf("data/ak_regions_simp.shp")

```


Simple plot

```{r}
plot(ak_region)
```

Exploratory analysis

```{r}
class(ak_region)

st_crs(ak_region)
```

Transform crs

```{r transformingCRS}

ak_region_3338 <- ak_region %>% 
  st_transform(crs = 3338)

st_crs(ak_region_3338)


plot(ak_region_3338)
```

Projected data can be plotted in 2D.

Unprojected assumes its 3D object.

## sf and tidyverse


```{r}
pop <- read.csv("data/alaska_population.csv")
```

```{r}
pop_4326 <- st_as_sf(pop,
                     coords = c("lng", "lat"),
                     crs = 4326,
                     remove = FALSE)

head(pop_4326)
```


## Assing each city to a region using `st_join()`

```{r}
# pop_joind <- st_join(pop_4326, ak_region_3338, join = st_within)
```

In order to be able to join we need both sf objects to have the same projection.

## transfroming pop data
```{r}
pop_3338 <- st_transform(pop_4326, crs = 3338)
```

Now join data


```{r}
pop_joind <- st_join(pop_3338, ak_region_3338, join = st_within)
```


## Calculate the total population for a region
```{r}
pop_region <- pop_joind %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarise(total_pop = sum(population))

```


## Adding geometry

```{r}
pop_region_3338 <- ak_region_3338 %>% 
  left_join(pop_region, by = "region")


plot(pop_region_3338["total_pop"])
```


## Grouping by mgmt areas
```{r}
pop_mgmt_3338 <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarise(total_pop = sum(total_pop),
            do_union = FALSE)

plot(pop_mgmt_3338["total_pop"])
```


## Writting special data using `write_sf()`

```{r}
write_sf(pop_region_3338, "data/ak_regions_population.shp", delete_layer = TRUE)

##delete_layer = TURE allows you to override your shp files when ever you run write_sf again.
```


## Plotting shp files with `ggplot2`

```{r}
ggplot(pop_region_3338)+
  geom_sf(aes(fill = total_pop))+
  theme_bw()+
  labs( fill = "Total Population")+
  scale_fill_continuous(low = "khaki", high = "firebrick", labels = comma)
  
```





